---
title: "Lab 5"
author: "Jean-Baptiste Astruc"
format:
  html: default
  pdf: default
echo: false
---

```{r Setup, include = FALSE}
here::i_am("Lab 5.Rproj")
library(here)
library(dplyr)
library(ggplot2)
library(tidyr)
theme_set(theme_bw())
```

## Exercise 1

```{r}
spotify <- readRDS(here("top-spotify-hits-2000-2019.Rds"))
```

### Question 1

```{r}
spotify_songs <- spotify |> 
  distinct(artist, song, .keep_all = TRUE)

nbsongs <- nrow(spotify_songs)
```

The data set contains `r nbsongs` songs.

```{r}
spotify_artists <- spotify |>
  distinct(artist)

nbartists <- nrow(spotify_artists)
```

The data set contains `r nbartists` artists.

```{r}
spotify_genre <- spotify |>
  distinct(genre)

nbgenre <- nrow(spotify_genre)
```

The data set contains `r nbgenre` genres.

### Question 2

```{r}
artists_by_year <- spotify |>
  group_by(year) |>
  distinct(artist)|>
  summarise(nbartists_by_year = n())

table1 <- knitr::kable(artists_by_year, col.names = c("Year", "Number of distinct artists per year"))
```

Here is the table: 

`r table1`

### Question 3

```{r}
most_pop_artist_table <- spotify_songs |>
  group_by(artist) |>
  distinct(song) |>
  summarise(nbsongs_by_year = n()) |>
  arrange(desc(nbsongs_by_year)) |>
  slice_head(n = 1)

most_pop_artist_name <- most_pop_artist_table |>
  pull(artist)

most_pop_artist_nbsongs <- most_pop_artist_table |>
  pull(nbsongs_by_year)
```

The most popular artist is `r most_pop_artist_name` with `r most_pop_artist_nbsongs` songs.

### Question 4

```{r}
tempo_genre <- spotify_songs |>
  group_by(genre) |>
  distinct(tempo, song, .keep_all = TRUE) |>
  summarise(min(tempo), max(tempo), mean(tempo), median(tempo), n())

table2 <- knitr::kable(tempo_genre)
```

Here is the table:

`r table2`

### Question 5

```{r}
liveness_danceability <- spotify_songs |>
  group_by(year) |>
  distinct(liveness, danceability, .keep_all = TRUE) |>
  summarise(mean_liveness = mean(liveness), 
            mean_danceability = mean(danceability))
```


### Question 6

```{r}
ggplot(liveness_danceability, aes(x = year)) +
  geom_line(aes(y = mean_liveness, color = "Mean of the liveness")) +
  geom_line(aes(y = mean_danceability, color = "Mean of the danceability")) +
  xlab("Year") +
  ylab("Value of the mean") +
  ggtitle("Evolution of the mean of the danceability and the mean of the liveness over time")
```

## Exercise 2

```{r}
dropout <- readRDS(here("dropout.RDS"))
```

### Question 1

```{r}
median_admission_grade <- dropout |>
  group_by(Target, `Marital status`) |>
  summarise(median_ad = median(`Admission grade`), .groups = "drop_last")
```

### Question 2

```{r}
median_admission_grade_2 <- dropout |>
  group_by(Target, `Marital status`) |>
  summarise(median_ad = median(`Admission grade`), .groups = "drop_last") |>
  pivot_wider(values_from = median_ad, names_from = Target)

table3 <- knitr::kable(median_admission_grade_2)
```

Here is the table:

`r table3`

### Question 3

```{r}
curricular_units <- names(select(dropout, starts_with("Curricular units")))

median_curricular_units <- dropout |>
  group_by(Gender) |>
  summarize(across(all_of(curricular_units), ~median(., na.rm = TRUE)), .groups = 'drop')
```


### Question 4

```{r}
step1 <- dropout |>
  select(starts_with("Curricular units"), Gender) |>
  pivot_longer(cols = starts_with("Curricular units"),
                names_to = "Units",
                values_to = "Value")

final_table <- step1 |> 
  group_by(Units, Gender) |>
  summarize(Median_Value = median(Value, na.rm = TRUE), .groups = 'drop') |>
  pivot_wider(names_from = Gender,
              values_from = Median_Value)

table4 <- knitr::kable(final_table)
```

Here is the table:

`r table4`

